smooth-sea

### 2018-05-20 index.rmd
tidier panel computation, fixing section labels, avp forecast map as index right column

  #compute tract aggregates for CL listing count
  tract <- cl %>%
    collect %>% #bring db query into memory
    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
           GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent, seattle tracts
    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
    dplyr::select(listingDate, GEOID10, GISJOIN, seattle, matchAddress, matchType, cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
    mutate(listingDate = as.Date(listingDate),
           listingQtr = as.yearqtr(listingDate)) %>%
    filter(cleanBeds %in% c(0, 1, 2, 3)) %>%
    group_by(GEOID10, GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
    summarize(nListings = n(),
              n1B = sum(cleanBeds == 1),
              nGT2B = sum(cleanBeds > 2),
              medRent = median(cleanRent),
              med0B = median(cleanRent[cleanBeds==0]),
              med1B = median(cleanRent[cleanBeds==1]),
              med2B = median(cleanRent[cleanBeds==2]),
              med3B = median(cleanRent[cleanBeds==3]),
              lagRent = lag(cleanRent)) %>% #create tract aggregates
    mutate(pGT2B = nGT2B/nListings) %>%
    ungroup %>%
    arrange(GISJOIN, listingQtr)
  dbDisconnect(DB)
  write_csv(tract, "./input/tractCl.csv")

@@

census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]

census <- census %>%
  mutate(nHU = AF7PE001,
         pOwnoccHU = AF7PE002/AF7PE001,
         medHUVal = AF9LE001)

### Training periods

#studios
Q2_2017 <- tract %>%
  filter(listingQtr == "2017 Q2") %>%
  right_join(census) %>%
  ungroup %>%
  mutate(listingQtr = "2017 Q2",
         Qtr = "Q2",
         actualRent = med1B)

Q3_2017 <- tract %>%
  filter(listingQtr == "2017 Q3") %>%
  right_join(census) %>%
  ungroup %>%
  mutate(listingQtr = "2017 Q3",
         Qtr = "Q3",
         actualRent = med1B)

Q4_2017 <- tract %>%
  filter(listingQtr == "2017 Q4") %>%
  right_join(census) %>%
  ungroup %>%
  mutate(listingQtr = "2017 Q4",
         Qtr = "Q4",
         actualRent = med1B)

Q1_2018 <- tract %>%
  filter(listingQtr == "2018 Q1") %>%
  right_join(census) %>%
  ungroup %>%
  mutate(listingQtr = "2018 Q1",
         Qtr = "Q1",
         actualRent = med1B)

### Forecast/test period

Q2_2018 <- tract %>%
  filter(listingQtr == "2018 Q2") %>%
  right_join(census) %>%
  ungroup %>%
  mutate(listingQtr = "2018 Q2",
         Qtr = "Q2",
         actualRent = med1B,
         medRent = NA)
Q2_2018 <- Q2_2018[match(sea_shp@data$GISJOIN, Q2_2018$GISJOIN),]

sea_df <- bind_rows(Q2_2017, Q3_2017, Q4_2017, Q1_2018, Q2_2018)
sea_df$listingQtr <- as.yearqtr(sea_df$listingQtr)

sea_df <- sea_df %>% 
  dplyr::select(-GEOID10) %>%
  dplyr::select(GISJOIN, listingQtr, Qtr, medRent, med1B, lagRent, actualRent, nHU, 
                pOwnoccHU, medHUVal, pGT2B)

#### ACF for panel ------------------------------------------------------------

acf_df <- data.frame(sea_df = unique(sea_df$GISJOIN), acf = rep(NA, length(unique(sea_df$GISJOIN))))
for(i in unique(sea_df$GISJOIN)){
  obs <- sea_df[sea_df$GISJOIN == i,]
  ts <- ts(obs$medRent)
  if(sum(is.na(ts)) > 0){
    next
  } else{
    #print(ts)
    acf <- acf(ts)
    #print(acf)
    acf_df$acf[acf_df$sea_df == tract] = list(acf)
  }
}

#ACFs are weak with only 5 periods to assess with. Highest L1 value is around .5 but many are
#weakened by a noisey quarter

#### F. INLA Models of CL Rent ------------------------------------------------

@@

cens <- gCentroid(sea_shp, byid = T)
cens <- data.frame(long = cens@coords[,1],
                   lat = cens@coords[,2])
cens <- cens %>% mutate(row = row_number())
sea_ff <- fortify(sea_shp)

ggplot(sea_ff, aes(x = long, y = lat, group = group)) + 
  geom_polygon(fill = "grey90", color = "white") +
  geom_text(data = cens, aes(x  = long, y = lat, label = row, group = row))

sea_adj <- poly2nb(sea_shp)
nb2INLA("R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph", sea_adj)
sea_df$idtract <- rep(1:134, 5)
sea_df$idtract1 <- rep(1:134, 5)
sea_df$idqtr <- as.numeric(sea_df$listingQtr)
sea_df$idqtr1 <- sea_df$idqtr
sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$idqtr)

#Seattle median rent models -----------------------------------------------

@@

sea_df <- sea_df %>%
  mutate_at(.vars = vars(matches("_Med")),
            .funs = exp) %>%
  mutate_at(.vars = vars(matches("_SD")),
            .funs = exp) %>%
  group_by(GISJOIN) %>% 
  arrange(GISJOIN, listingQtr) %>%
  mutate(spt_traj = spt_Med - lag(spt_Med, 4)) %>%
  ungroup %>%
  group_by(GISJOIN) %>%
  mutate(spt_traj = sum(spt_traj, na.rm=T))

@@

```{r index.maps, echo=F}

smoothed <- ggplot(sea_f, 
                   aes(x = long, y = lat, group = group, fill = bym_Med, label = GISJOIN, frame = listingQtr)) +
  #facet_grid(~ listingQtr) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(limits = c(750, 2500), labels = scales::dollar) +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(title = "Spatially Smoothed 1B Median Rent",
       fill = "Smoothed")
ggplotly(smoothed, tooltip = c("label", "fill"))
```

>>>>

  #compute tract aggregates for CL listing count
  tract <- cl %>%
    collect %>% #bring db query into memory
    filter(!is.na(GISJOIN), !is.na(cleanBeds), !is.na(cleanRent), !is.na(cleanSqft), 
           GISJOIN %in% sea_shp@data$GISJOIN) %>% #only listings with valid Bed/Rent, seattle tracts
    distinct(cleanBeds, cleanRent, cleanSqft, matchAddress, 
             matchAddress2, .keep_all = T) %>% #dedupe to unique address-bed-rent combos
    dplyr::select(listingDate, GISJOIN, seattle, matchAddress, matchType, 
                  cleanBeds, cleanRent, cleanSqft) %>% #SELECT these columns
    mutate(listingDate = as.Date(listingDate),
           listingQtr = as.yearqtr(listingDate)) %>%
    filter(cleanBeds %in% c(0, 1, 2, 3), listingQtr >= "2017 Q2") %>%
    group_by(GISJOIN, listingQtr) %>% #group listings by tract, qtr within tract
    summarize(nListings = n(),
              n1B = sum(cleanBeds == 1),
              nGT2B = sum(cleanBeds > 2),
              medRent = median(cleanRent),
              med0B = median(cleanRent[cleanBeds==0]),
              med1B = median(cleanRent[cleanBeds==1]),
              med2B = median(cleanRent[cleanBeds==2]),
              med3B = median(cleanRent[cleanBeds==3]),
              lagRent = lag(cleanRent)) %>% #create tract aggregates
    mutate(pGT2B = nGT2B/nListings) %>%
    ungroup %>%
    arrange(GISJOIN, listingQtr)
  dbDisconnect(DB)
  write_csv(tract, "./input/tractCl.csv")

@@

#ensure order of census df matches shapefile
census <- census[match(sea_shp@data$GISJOIN, census$GISJOIN),]

#clean a few fields we might use
census <- census %>%
  mutate(nHU = AF7PE001,
         pOwnoccHU = AF7PE002/AF7PE001,
         medHUVal = AF9LE001)

#store NULL object to bind iterated dfs into
panel <- NULL

#for each unique quarter in the CL data
for(i in unique(as.character(tract$listingQtr))){

  period <- tract %>% #start pipe with tract, end with period
    mutate(listingQtr = factor(listingQtr, ordered = T)) %>% #make factor
    filter(listingQtr == i) %>% #filter to qtr i
    right_join(census) %>% #right join to census, allows missingness
    ungroup() %>% #remove grouping by tract
    mutate(listingQtr = i, #listingQtr == current iteration
           Qtr = gsub(pattern = "^\\d{4} ", replacement = "", x = listingQtr), #subset to QX
           actualRent = med1B) #actual for actual versus predicted
  panel <- bind_rows(panel, period) #append to panel object
}

sea_df <- panel %>% #set to NA for INLA to forecast
  mutate(med1B = ifelse(listingQtr == as.yearqtr(Sys.Date()), NA, med1B))%>%
  dplyr::select(GISJOIN, listingQtr, Qtr, medRent, med1B, lagRent, actualRent, nHU, n1B,
                pOwnoccHU, medHUVal, pGT2B) %>%
  arrange(listingQtr, GISJOIN)

#### B. INLA Models of CL Rent ------------------------------------------------

@@

#compute centroid for map of tract numbers
cens <- gCentroid(sea_shp, byid = T)
cens <- data.frame(long = cens@coords[,1],
                   lat = cens@coords[,2])
cens <- cens %>% mutate(row = row_number())
sea_ff <- fortify(sea_shp)

#plot seattle tracts with tract number labels
ggplot(sea_ff, aes(x = long, y = lat, group = group)) + 
  geom_polygon(fill = "grey90", color = "white") +
  geom_text(data = cens, aes(x  = long, y = lat, label = row, group = row))

#create adjacency matrix from shapefile
sea_adj <- poly2nb(sea_shp)

#create neighbor file that INLA takes for hyperparameter args
nb2INLA("R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph", sea_adj)
sea_df$idtract <- rep(1:134, 5) #a tract ID (nb: panel must be ordered tract)
sea_df$idtract1 <- rep(1:134, 5)
sea_df$idqtr <- factor(sea_df$listingQtr, ordered = T)
sea_df$idqtr1 <- sea_df$idqtr
sea_df$idtractqtr <- paste(sea_df$idtract, sea_df$idqtr)

@@

sea_df <- sea_df %>%
  mutate_at(.vars = vars(matches("_Med")),
            .funs = exp) %>%
  mutate_at(.vars = vars(matches("_SD")),
            .funs = exp) %>%
  group_by(GISJOIN) %>% 
  arrange(GISJOIN, listingQtr) %>%
  mutate(spt_traj = spt_Med - lag(spt_Med, 4)) %>%
  ungroup %>%
  group_by(GISJOIN) %>%
  mutate(spt_traj = sum(spt_traj, na.rm=T)) %>%
  ungroup() %>%
  mutate(bym_cuts = cut(bym_Med, 6),
         spt_cuts = cut(spt_Med, 6))

@@

```{r index.maps, echo=F}
obs <- sea_f %>%
  dplyr::select(long, lat, group, GISJOIN, listingQtr, n1B, rent = actualRent) %>%
  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
  mutate(var = "Observed")

pred <- sea_f %>%
  dplyr::select(long, lat, group, GISJOIN, listingQtr, n1B, rent = nsar1_Med) %>%
  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
  mutate(var = "Forecasted")

comb <- bind_rows(obs, pred)

avp.gg <- ggplot(comb, aes(x = long, y = lat, group = group, fill = rent, label = GISJOIN, text = n1B)) +
  facet_wrap(~ var) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(limits = c(750, 2500), labels = scales::dollar) +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median 1B Rent")

ggplotly(avp.gg, tooltip = c("facet", "label", "fill", "text"), hovermode="compare") %>%
  animation_opts(frame = 1000, redraw = FALSE) 
```
